---
title: "Testing sismonr"
author: "Olivia Angelin-Bonnet"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(sismonr)
library(tidyverse)
```

```{r ggplot-setup, cache = FALSE}
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))
theme_update(legend.position = "bottom")
```

## 2-genes activating system

We'll start by simulating a simple system with 2 genes: a transcription regulator, and its target; the regulator activates the transcription of the target. For the simulation, we'll generate one individual which has no mutations on the genes. We'll run 10 simulations for this individual, with a simulation time of 10,000 seconds (to make sure that we observe the effect of the regulation).

### Creating the system and individuals

```{r two-genes-system}
set.seed(364)

## Creating system with one regulator gene
my_system <- createInSilicoSystem(
  empty = TRUE, 
  G = 1,
  PC.p = 1,
  PC.TC.p = 1,
  TC.pos.p = 1,
  ploidy = 2
) |> 
  ## Adding the target gene
  addGene(
    coding = "PC",
    TargetReaction = "MR"
  ) |> 
  ## Adding the regulation of gene 2 by gene 1
  addEdge(
    regID = 1,
    tarID = 2,
    regsign = "1"
  )
```

```{r show-system}
my_system$genes
my_system$edg
```

```{r create-pop}
my_pop <- createInSilicoPopulation(
  1,
  my_system,
  ngenevariants = 1
)
```


### Running a simulation

```{r run-sim-null}
set.seed(948)
sim_null <- simulateInSilicoSystem(
  my_system,
  my_pop,
  simtime = 10000,
  ntrials = 10
)
```

```{r show-sim-null}
plotSimulation(sim_null$Simulation)
```

```{r save-sim-res}
df_null <- sim_null$Simulation |> 
  mergeAlleleAbundance() |> 
  as_tibble() |> 
  mutate(data_type = "original")
```


## Silencing target gene

Now we'll modify the system by silencing the transcription of the target gene. We do this by setting its transcripton rate to a very small value. We didn't modify any property of the regulator or of the regulation.

```{r system-silencing}
my_system_silencing <- my_system
my_pop_silencing <- my_pop

my_system_silencing$genes[2, "TCrate"] <-  1e-20
```


```{r show-system-silencing}
my_system_silencing$genes
my_system_silencing$edg
```


```{r run-sim-silencing}
set.seed(948)
sim_silencing <- simulateInSilicoSystem(
  my_system_silencing,
  my_pop_silencing,
  simtime = 10000,
  ntrials = 10
)
```

```{r show-sim-silencing}
plotSimulation(sim_silencing$Simulation)
```

```{r save-sim-silencing}
df_silencing <- sim_silencing$Simulation |> 
  mergeAlleleAbundance() |> 
  as_tibble() |> 
  mutate(data_type = "silencing")
```

## Comparing original and silencing simulations

The plot below compares the RNA and protein abundance of each gene between the two simulations. As expected, only the expression of the target gene is impacted by the silencing.

```{r comparison-plot, fig.height = 8, fig.width = 10}
bind_rows(
  df_null,
  df_silencing
) |> 
  pivot_longer(
    cols = matches("(P|R)\\d"),
    names_to = "molecule",
    values_to = "abundance"
  ) |> 
  mutate(
    molecule = str_replace(molecule, "^P", "Protein abundance for gene "),
    molecule = str_replace(molecule, "^R", "RNA abundance for gene ")
  ) |> 
  ggplot(aes(x = time, y = abundance, colour = data_type)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~ molecule, scale = "free_y") +
  scale_colour_brewer(palette = "Set1") +
  labs(
    title = "Comparison of original and silencing simulations",
    x = "Time",
    y = "Abundance",
    colour = "Simulation "
  )
```




